Combinatorial Proofs and Decomposition Theorems for First-order Logic

### 2021-01-19 main.tex
Comments on IX and sequentialisation

\newcommand*{\LK}{\mathsf{LK}}
\newcommand*{\FOLK}{\mathsf{LK1}}
\newcommand*{\FOLKcut}{\FOLK\mathord+\cut}
\newcommand*{\LKm}{\mathsf{LK + mix}}
\newcommand*{\MLL}{\mathsf{MLL}}
\newcommand*{\FOMLL}{\mathsf{MLL1^X}} 
\newcommand*{\FOMLLcut}{\FOMLL\mathord+\cut}
\newcommand*{\FOMLLm}{\mathsf{MLL1+mix}}
\newcommand*{\FOKS}{\mathsf{KS1}}
\newcommand*{\FOMLS}{\mathsf{MLS1^X}}

@@

\begin{thm}
  Any unification net can be obtained via the translation $\sqntl\cdot$ given in Theorem
  \ref{thm:MLL1->Unet}.
\end{thm}
Let $\lambda$ be a unification net on $\Gamma$. We proceed by induction on the
number of connectives in $\Gamma$. In the base case, $\Gamma$ is of the form
\begin{align*}
p_1(t_{11}, \ldots, t_{1n_1}), \dual{p_1}(t_{11}, \ldots, t_{1n_1}), \ldots, \\
p_k(t_{k1}, \ldots, t_{kn_k}), \dual{p_k}(t_{k1}, \ldots, t_{kn_k}),
\underbrace{\ttt, \ldots, \ttt}_\text{$m$ times}
\end{align*} and
$\lambda$ is the linking $\set{(a_1, \cneg{a_1}), \ldots, (a_k, \cneg{a_k})}$,
where $a_i = p_i(t_{i1}, \ldots, t_{in_i})$, which equals to $\sqntl\Pi$, where
$\Pi$ is the proof consisting of $m$ instances of the $\ttt$ rule, $n$ instances
$\vlinf{\ax}{}{\sqn a_i, \cneg{a_i}}{}$ of the $\ax$ rule, and followed by
$m+k-1$ instances of the $\mix$ rule.

Now we consider the inductive cases:
\begin{itemize}
  \item $\Gamma = \Delta, A \vlor B$:
    Let $\Gamma' = \Delta, A, B$. Define $\lambda'$ on $\Gamma'$ using the same
    links as $\lambda$ by identifying the leaves of $\formtree{\Gamma'}$ with those
    of $\formtree{\Gamma}$. We now check that $\lambda'$ is a unification net:
    \begin{itemize}
      \item The most general dualizer of $\lambda$ is also the most general
dualizer of $\lambda'$ as they correspond to the same unification problem.
Hence, the unification structure $\unifstr{\lambda'}$ is equal to the
restriction of $\unifstr{\lambda}$ to the nodes of $\formtree{\Gamma'}$.
      \item Every switching graph of $\lambda'$ is acyclic: if there were some
switching graph of $\unifstr{\lambda'}$ containing a cycle, it would induce a switching
graph of $\unifstr{\lambda}$ containing also a cycle, by adding an edge from the root of
$\formtree{A}$ to the $\vlor$ node in question.
    \end{itemize}
  \item $\Gamma = \Delta, \forall x.A$:
    Let $\Gamma' = \Delta, A$. Define $\lambda'$ on $\Gamma'$ using the same
links as $\lambda$. We now check that $\lambda'$ is a unification net:
  \begin{itemize}
    \item The most general dualizer of $\lambda$ is also the most general
dualizer of $\lambda'$ as they correspond to the same unification problem.
    \item Every switching graph of $\unifstr{\lambda'}$ is acyclic: if there
were some switching graph of $\unifstr{\lambda'}$ containing a cycle, it would
induce a switching graph of $\unifstr{\lambda}$ contaiting also a cycle, by
adding an edge from the root of $\formtree{A}$ to the $\forall$ node in
question.
  \end{itemize}
  \item $\formtree{\Gamma}$ has a root $\exists x$ with no outgoing dependency
edge:
\end{itemize}

@@

  always knowing that we can add instances of $\fequ$ as needed. To

@@

@@

        \vlinf{\mfaD}{}{\Scons{\forall x.F}}{S\cons{\forall x.\forall x.F}}

@@

>>>>

\newcommand*{\LK}{\mathsf{LK}}
\newcommand*{\FOLK}{\mathsf{LK1}}
\newcommand*{\FOLKcut}{\FOLK\mathord+\cut}
\newcommand*{\LKm}{\mathsf{LK + mix}}
\newcommand*{\MLL}{\mathsf{MLL}}
\newcommand*{\MLLm}{\mathsf{MLL+mix}}
\newcommand*{\FOMLL}{\mathsf{MLL1^X}} 
\newcommand*{\FOMLLcut}{\FOMLL\mathord+\cut}
\newcommand*{\FOMLLm}{\mathsf{MLL1+mix}}
\newcommand*{\FOKS}{\mathsf{KS1}}
\newcommand*{\FOMLS}{\mathsf{MLS1^X}}

@@

\begin{thm}
  Any unification net can be obtained via the translation $\sqntl\cdot$ given in Theorem
  \ref{thm:MLL1->Unet}.
\end{thm}

\begin{lemma}\label{lem:concomp}
  The number of connected components of an acylic graph $\gG = \tuple{\vG, \eG}$ is
equal to $|\eG| - |\vG|$.
\end{lemma}

\begin{proof}
  By a straightforward induction on $|\vG|$.
\end{proof}

\begin{lemma}
  The number of connected components is the same for any switching graph of a
unification net.
\end{lemma}

\begin{proof}
  An immediate consequence of Lemma \ref{lem:concomp}.
\end{proof}

In the proof, we also use the notion of \bfit{frame} introduced by Hughes in
\cite{hughes:unifn}.

\begin{definition}
Let $\lambda$ be a unification net on an $\FOMLL$ sequent $\Gamma$.
We define the \bfit{frame} of $\lambda$ by exhaustively applying the following
subformula rewriting steps, to obtain a linking $\lambda_m$ on an $\MLLm$ sequent $\Gamma_m$:

\begin{enumerate}
  \item {\bf Encode dependencies as fresh links.} For each dependency $\exists x
	  \rightarrow \forall y$, with corresponding subformulas $\exists x A$
		and $\forall y B$, we add a fresh link as follows. Let $P$ be a
		fresh (nullary) predicate symbol. Replace $\exists x A$ with $P
		\cand \exists x A$ and $\forall y B$ with $\overline{P} \cor \forall y B$, and add an axiom link between $P$ and $\overline{P}$.
  \item {\bf Erase quantifiers.} After step 1, erase all the quantifiers. (We no longer need their leaps since they are encoded as links in step 1.
  \item {\bf Simplify atoms.} After step 2, replace every predicate $Pt_1 \cdots t_n$ with a nullary predicate symbol $P$.
\end{enumerate}
\end{definition}
Note that the linking $\lambda_m$ is a valid $\MLLm$ proof net.

\begin{lemma}\label{lem:concomp_switching}
Suppose that $\lambda$ is a $\MLLm$ proof net which is connected and such that any of its
switching graphs is not connected. Then there exists a $\vlor$ node in
$\unifstr{\lambda}$ such that $\lambda$ is correct on the sequent $\Gamma'$
obtained from $\Gamma$ by replacing this $\vlor$ by a $\vlan$.
\end{lemma}

\begin{lemma}
Suppose that $\lambda$ is a $\FOMLL$ proof net which is connected and such that any of its
switching graphs is not connected. Then there exists a $\vlor$ node in
$\unifstr{\lambda}$ such that $\lambda$ is correct on the sequent $\Gamma'$
obtained from $\Gamma$ by replacing this $\vlor$ by a $\vlan$.
\end{lemma}

\begin{proof}
  Consider the frame $\lambda_m$ of $\lambda$. The number of any switching
graph of $\unifstr{\lambda}$ is equal to that of $\unifstr{\lambda_m}$. Apply
Lemma \ref{lem:concomp_switching} and it is clear that such $\vlor$ cannot be one
of the fresh $\vlor$'s added during the frame construction.
\todo{}
\end{proof}

\begin{proof}\label{lem:concomp_switching_fo}
Let $\lambda$ be a unification net on $\Gamma$. We proceed by induction on the
number of connected components of the unification structure $\unifstr{\lambda}$:
\begin{itemize}
  \item If there is only one connected component, we proceed by induction on the
number $k$ of connected components of any switching graph of
$\unifstr{\lambda}$. If $k = 1$, we obtain a proof $\Pi$ in $\FOMLL$ such that
$\sqntl{\Pi} = \lambda$ by applying \cite[Theorem~3]{hughes:unifn}. If $k > 1$,
using the Lemma \ref{lem:concomp_switching_fo}, we obtain a sequent $\Gamma'$ on
which $\lambda$ is correct by transforming a $\vlor$ node into a $\vlan$. By
induction hypothesis, there is a proof $\Pi'$ in $\FOMLL$ whose translation is
$\lambda$. By considering the $\vlan$ rule instance corresponding to the $\vlan$
node in $\Pi'$, we have:
$
\Pi' =
\vlderivation{
  \vlin{}{}{\sqns{\Gamma'}}{
  \vlin{}{}{\vlvdots}{
    \vliin{\vlan}{}{\sqns{\Delta_1, A \vlan B, \Delta_2}}{
      \vltrl{\Pi_1}{}{}{\sqns{\Delta_1, A}}{
        \vlhy{\quad}}{
        \vlhy{}}{
        \vlhy{}}}{
      \vltrl{\Pi_2}{}{}{\sqns{B, \Delta_2}}{
        \vlhy{\quad}}{
        \vlhy{}}{
        \vlhy{}}}
}}}
$.
We can thus obtain a proof $\Pi$ of $\Gamma$:
$
\Pi =
\vlderivation{
  \vlin{}{}{\sqns{\Gamma}}{
  \vlin{}{}{\vlvdots}{
    \vlin{\vlor}{}{\sqns{\Delta_1, A \vlor B, \Delta_2}}{
    \vliin{\mix}{}{\sqns{\Delta_1, A, B, \Delta_2}}{
      \vltrl{\Pi_1}{}{}{\sqns{\Delta_1, A}}{
        \vlhy{\quad}}{
        \vlhy{}}{
        \vlhy{}}}{
      \vltrl{\Pi_2}{}{}{\sqns{B, \Delta_2}}{
        \vlhy{\quad}}{
        \vlhy{}}{
        \vlhy{}}}
}}}}
$.

\todo{}
\end{itemize}

\iffalse
We proceed by induction on the
number of connectives in $\Gamma$. In the base case, $\Gamma$ is of the form
\begin{align*}
p_1(t_{11}, \ldots, t_{1n_1}), \dual{p_1}(t_{11}, \ldots, t_{1n_1}), \ldots, \\
p_k(t_{k1}, \ldots, t_{kn_k}), \dual{p_k}(t_{k1}, \ldots, t_{kn_k}),
\underbrace{\ttt, \ldots, \ttt}_\text{$m$ times}
\end{align*} and
$\lambda$ is the linking $\set{(a_1, \cneg{a_1}), \ldots, (a_k, \cneg{a_k})}$,
where $a_i = p_i(t_{i1}, \ldots, t_{in_i})$, which equals to $\sqntl\Pi$, where
$\Pi$ is the proof consisting of $m$ instances of the $\ttt$ rule, $n$ instances
$\vlinf{\ax}{}{\sqn a_i, \cneg{a_i}}{}$ of the $\ax$ rule, and followed by
$m+k-1$ instances of the $\mix$ rule.

Now we consider the inductive cases:
\begin{itemize}
  \item $\Gamma = \Delta, A \vlor B$:
    Let $\Gamma' = \Delta, A, B$. Define $\lambda'$ on $\Gamma'$ using the same
    links as $\lambda$ by identifying the leaves of $\formtree{\Gamma'}$ with those
    of $\formtree{\Gamma}$. We now check that $\lambda'$ is a unification net:
    \begin{itemize}
      \item The most general dualizer of $\lambda$ is also the most general
dualizer of $\lambda'$ as they correspond to the same unification problem.
Hence, the unification structure $\unifstr{\lambda'}$ is equal to the
restriction of $\unifstr{\lambda}$ to the nodes of $\formtree{\Gamma'}$.
      \item Every switching graph of $\lambda'$ is acyclic: if there were some
switching graph of $\unifstr{\lambda'}$ containing a cycle, it would induce a switching
graph of $\unifstr{\lambda}$ containing also a cycle, by adding an edge from the root of
$\formtree{A}$ to the $\vlor$ node in question.
    \end{itemize}
  \item $\Gamma = \Delta, \forall x.A$:
    Let $\Gamma' = \Delta, A$. Define $\lambda'$ on $\Gamma'$ using the same
links as $\lambda$. We now check that $\lambda'$ is a unification net:
  \begin{itemize}
    \item The most general dualizer of $\lambda$ is also the most general
dualizer of $\lambda'$ as they correspond to the same unification problem.
    \item Every switching graph of $\unifstr{\lambda'}$ is acyclic: if there
were some switching graph of $\unifstr{\lambda'}$ containing a cycle, it would
induce a switching graph of $\unifstr{\lambda}$ contaiting also a cycle, by
adding an edge from the root of $\formtree{A}$ to the $\forall$ node in
question.
  \end{itemize}
  \item $\formtree{\Gamma}$ has a root $\exists x$ with no outgoing dependency
edge:
\end{itemize}
\fi
\end{proof}

@@

  always knowing that we can add instances of $\fequ$ as needed. \juihsuan{this
does not seem clear to me. What if from $\PE A_1$ to $\PE B_1$ there are just a
bunch of $\PE\fequ$? What do we do in this case?} To

@@

    \juihsuan{We should say that $\Sconhole, E, $ and $F$ are also propositional
encodings. Here I think you kind of identify a formula with its prop. encoding.}

@@

        \vlinf{\cfaD}{}{\Scons{\forall x.F}}{S\cons{\forall x.\forall x.F}}

@@

    \juihsuan{maybe need some words to exclude the case in which $C$ (or $D$)
is a propositional variable.}
